\chapter*{Acknowledgements}

\begin{small}
I would like to thank the following people, in no particular order:

My supervisor, Ing. Josef Kortan, for his willingness to supervise my work and undergo the formalities surrounding it, and his continued supply of encouragement.

My previous supervisor, Ing. David Sedláček, Ph.D., for his willingness to let me explore a variety of topics under his supervision, while I was still uncertain about the topic of my master's thesis.

The Department of Computer Graphics and Interaction at FEL CTE, and Ing. David Sedláček, Ph.D. in particular, for making the VR laboratory accessible for the development of the application and for the evaluation of the application, during which I conducted a study in the VR laboratory over the course of 2 weeks.

Mgr. et Mgr. Iveta Fajnerová, Ph.D., for her invaluable input during the initial stage of the project and her willingness to mentor me through the process of designing the study for the evaluation of the developed application.

The developers of the \textit{iTrip} application, and MUDr. Filip Tylš, Ph.D. in particular, for providing the czech translation of the psychometric questionnaire.

Dr. Timo Torsten Schmidt, for his mentorship regarding the scoring of the psychometric questionnaires.

Ossi Luoto, Fabrice Piquet, Jason Hoku, and Ing. Ondřej Slabý, for their invaluable advice regarding development in Unreal Engine 4.

Daniel McKenna, Samuel Lim, Max Fladung, and ``Fox'', for offering their time to proofread this document.

The generous participants of the study, who were offered no financial or material compensation for their time.

Alexandra Elbakyan for founding and her continued maintenance of the Sci-Hub website, without which this endeavour would have most definitely been impossible. After all, \textit{we are standing on the shoulders of giants}.

Finally, I would like to thank my parents for their continued support throughout my pursuit of advanced studies.
\end{small}

% \vspace*{1cm}
\begin{flushright}
---Jakub Hlusička
\end{flushright}

\vfill
\begin{center}
{\large
    \textit{``Not everything that is faced can be changed,\hspace{40pt}\\
    \hspace{40pt} but nothing can be changed until it is faced.''}
}
\end{center}
\begin{flushright}
    ---James Baldwin, As Much Truth as One Can Bear (1962)
\end{flushright}
